---
title: "Doodling ROC & Confusion Matrix"
author: "[Sander W. van der Laan, PhD](https://swvanderlaan.github.io) | @swvanderlaan | s.w.vanderlaan@gmail.com"
date: "`r Sys.Date()`"
output:
  html_notebook: 
    cache: yes
    code_folding: hide
    collapse: yes
    df_print: paged
    fig.align: center
    fig_caption: yes
    fig_width: 12
    fig_height: 8
    fig_retina: 2
    highlight: default #tango kate haddock
    theme: spacelab #cosmo lumen paper yeti
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
mainfont: Arial
subtitle: ""
editor_options:
  chunk_output_type: inline
---

```{r global_options, include = FALSE}
# further define some knitr-options.
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, fig.path = 'Figures/', 
                      warning = TRUE, # show warnings during codebook generation
                      message = TRUE, # show messages during codebook generation
                      error = TRUE, # do not interrupt codebook generation in case of errors; usually better for debugging
                      echo = TRUE,  # show R code
                      eval = TRUE)
ggplot2::theme_set(ggplot2::theme_minimal())
pander::panderOptions("table.split.table", Inf)
```

# Setup
We will clean the environment, setup the locations, define colors, and create a datestamp.

_Clean the environment._
```{r echo = FALSE}
rm(list = ls())
```

_Set locations and working directories..._
```{r LocalSystem, echo = FALSE}
### Operating System Version

### MacBook
ROOT_loc = "/Users/swvanderlaan"
# GENOMIC_loc = paste0(ROOT_loc, "/iCloud/Genomics")
# LAB_loc = paste0(GENOMIC_loc, "/LabBusiness")
PLINK_loc = paste0(ROOT_loc,"/PLINK")
COVID19_loc = paste0(PLINK_loc,"/analyses/covid19_hgi")
PROJECT_loc = paste0(COVID19_loc,"/covid19_hg")
SCRIPTS_loc = paste0(PROJECT_loc,"/SCRIPTS")
# RAWDATA_loc = paste0(COVID19_loc,"/RESULTS/COVID19_HGI_ANA5_20200429") # first release
RAWDATA_loc = paste0(COVID19_loc,"/RESULTS/COVID19_HGI_ANA5_20200508") # second release

### SOME VARIABLES WE NEED DOWN THE LINE
# TRAIT_OF_INTEREST = "UCORBIO" # Phenotype
PROJECTNAME = "COVID19" # primary goal of this analysis
SUBPROJECTNAME = "PRELIMINARYRELEASE2" # e.g. study name

cat("\nCreate a new analysis directory...\n")
ifelse(!dir.exists(file.path(PROJECT_loc, "/",paste0(PROJECTNAME,".",SUBPROJECTNAME))), 
       dir.create(file.path(PROJECT_loc, "/",paste0(PROJECTNAME,".",SUBPROJECTNAME))), 
       FALSE)
ANALYSIS_loc = paste0(PROJECT_loc,"/",paste0(PROJECTNAME,".",SUBPROJECTNAME))
ifelse(!dir.exists(file.path(ANALYSIS_loc, "/PLOTS")), 
       dir.create(file.path(ANALYSIS_loc, "/PLOTS")), 
       FALSE)
PLOT_loc = paste0(ANALYSIS_loc,"/PLOTS")
ifelse(!dir.exists(file.path(PLOT_loc, "/QC")), 
       dir.create(file.path(PLOT_loc, "/QC")), 
       FALSE)
QC_loc = paste0(PLOT_loc,"/QC")
ifelse(!dir.exists(file.path(ANALYSIS_loc, "/OUTPUT")), 
       dir.create(file.path(ANALYSIS_loc, "/OUTPUT")), 
       FALSE)
OUT_loc = paste0(ANALYSIS_loc, "/OUTPUT")
ifelse(!dir.exists(file.path(ANALYSIS_loc, "/BASELINE")), 
       dir.create(file.path(ANALYSIS_loc, "/BASELINE")), 
       FALSE)
BASELINE_loc = paste0(ANALYSIS_loc, "/BASELINE")
setwd(paste0(PROJECT_loc))
getwd()
list.files()
```

_... a package-installation function ..._
```{r Function: installations}
install.packages.auto <- function(x) { 
  x <- as.character(substitute(x)) 
  if (isTRUE(x %in% .packages(all.available = TRUE))) { 
    eval(parse(text = sprintf("require(\"%s\")", x)))
  } else { 
    # Update installed packages - this may mean a full upgrade of R, which in turn
    # may not be warrented. 
    #update.packages(ask = FALSE) 
    eval(parse(text = sprintf("install.packages(\"%s\", dependencies = TRUE, repos = \"http://cran-mirror.cs.uu.nl/\")", x)))
  }
  if (isTRUE(x %in% .packages(all.available = TRUE))) { 
    eval(parse(text = sprintf("require(\"%s\")", x)))
  } else {
    source("http://bioconductor.org/biocLite.R")
    # Update installed packages - this may mean a full upgrade of R, which in turn
    # may not be warrented.
    #biocLite(character(), ask = FALSE) 
    eval(parse(text = sprintf("biocLite(\"%s\")", x)))
    eval(parse(text = sprintf("require(\"%s\")", x)))
  }
}
```

_... and load those packages._
```{r loading_packages}
install.packages.auto("readr")
install.packages.auto("optparse")
install.packages.auto("tools")
install.packages.auto("dplyr")
install.packages.auto("tidyr")
install.packages.auto("naniar")
# To get 'data.table' with 'fwrite' to be able to directly write gzipped-files
# Ref: https://stackoverflow.com/questions/42788401/is-possible-to-use-fwrite-from-data-table-with-gzfile
# install.packages("data.table", repos = "https://Rdatatable.gitlab.io/data.table")
library(data.table)
install.packages.auto("tidyverse")
install.packages.auto("knitr")
install.packages.auto("DT")
install.packages.auto("eeptools")
install.packages.auto("haven")
install.packages.auto("tableone")
install.packages.auto("ggpubr")
install.packages.auto("BlandAltmanLeh")
install.packages.auto("ggplot2")
install.packages.auto("CMplot")
# For regional association plots
devtools::install_github("jrs95/gassocplot")
```


_We will create a datestamp and define the Utrecht Science Park Colour Scheme_.
```{r Setting: Colors}
Today = format(as.Date(as.POSIXlt(Sys.time())), "%Y%m%d")
Today.Report = format(as.Date(as.POSIXlt(Sys.time())), "%A, %B %d, %Y")
### UtrechtScienceParkColoursScheme
###
### WebsitetoconvertHEXtoRGB:http://hex.colorrrs.com.
### Forsomefunctionsyoushoulddividethesenumbersby255.
###
###	No.	Color			      HEX	(RGB)						              CHR		  MAF/INFO
###---------------------------------------------------------------------------------------
###	1	  yellow			    #FBB820 (251,184,32)				      =>	1		or 1.0>INFO
###	2	  gold			      #F59D10 (245,157,16)				      =>	2		
###	3	  salmon			    #E55738 (229,87,56)				      =>	3		or 0.05<MAF<0.2 or 0.4<INFO<0.6
###	4	  darkpink		    #DB003F ((219,0,63)				      =>	4		
###	5	  lightpink		    #E35493 (227,84,147)				      =>	5		or 0.8<INFO<1.0
###	6	  pink			      #D5267B (213,38,123)				      =>	6		
###	7	  hardpink		    #CC0071 (204,0,113)				      =>	7		
###	8	  lightpurple	    #A8448A (168,68,138)				      =>	8		
###	9	  purple			    #9A3480 (154,52,128)				      =>	9		
###	10	lavendel		    #8D5B9A (141,91,154)				      =>	10		
###	11	bluepurple		  #705296 (112,82,150)				      =>	11		
###	12	purpleblue		  #686AA9 (104,106,169)			      =>	12		
###	13	lightpurpleblue	#6173AD (97,115,173/101,120,180)	=>	13		
###	14	seablue			    #4C81BF (76,129,191)				      =>	14		
###	15	skyblue			    #2F8BC9 (47,139,201)				      =>	15		
###	16	azurblue		    #1290D9 (18,144,217)				      =>	16		or 0.01<MAF<0.05 or 0.2<INFO<0.4
###	17	lightazurblue	  #1396D8 (19,150,216)				      =>	17		
###	18	greenblue		    #15A6C1 (21,166,193)				      =>	18		
###	19	seaweedgreen	  #5EB17F (94,177,127)				      =>	19		
###	20	yellowgreen		  #86B833 (134,184,51)				      =>	20		
###	21	lightmossgreen	#C5D220 (197,210,32)				      =>	21		
###	22	mossgreen		    #9FC228 (159,194,40)				      =>	22		or MAF>0.20 or 0.6<INFO<0.8
###	23	lightgreen	  	#78B113 (120,177,19)				      =>	23/X
###	24	green			      #49A01D (73,160,29)				      =>	24/Y
###	25	grey			      #595A5C (89,90,92)				        =>	25/XY	or MAF<0.01 or 0.0<INFO<0.2
###	26	lightgrey		    #A2A3A4	(162,163,164)			      =>	26/MT
###
###	ADDITIONAL COLORS
###	27	midgrey			#D7D8D7
###	28	verylightgrey	#ECECEC"
###	29	white			#FFFFFF
###	30	black			#000000
###----------------------------------------------------------------------------------------------
uithof_color = c("#FBB820","#F59D10","#E55738","#DB003F","#E35493","#D5267B",
                 "#CC0071","#A8448A","#9A3480","#8D5B9A","#705296","#686AA9",
                 "#6173AD","#4C81BF","#2F8BC9","#1290D9","#1396D8","#15A6C1",
                 "#5EB17F","#86B833","#C5D220","#9FC228","#78B113","#49A01D",
                 "#595A5C","#A2A3A4", "#D7D8D7", "#ECECEC", "#FFFFFF", "#000000")
uithof_color_legend = c("#FBB820", "#F59D10", "#E55738", "#DB003F", "#E35493",
                        "#D5267B", "#CC0071", "#A8448A", "#9A3480", "#8D5B9A",
                        "#705296", "#686AA9", "#6173AD", "#4C81BF", "#2F8BC9",
                        "#1290D9", "#1396D8", "#15A6C1", "#5EB17F", "#86B833",
                        "#C5D220", "#9FC228", "#78B113", "#49A01D", "#595A5C",
                        "#A2A3A4", "#D7D8D7", "#ECECEC", "#FFFFFF", "#000000")
### ----------------------------------------------------------------------------
```

In this repository

```{r Dataset}
# Reference: https://boostedml.com/2019/05/classification-accuracy-in-r-difference-between-accuracy-precision-recall-sensitivity-and-specificity.html
library(caret)
library(e1071)
library(pROC)
diabetes<-read.csv(url('https://raw.githubusercontent.com/jbrownlee/Datasets/master/pima-indians-diabetes.data.csv'))
N_all = nrow(diabetes)
N_train = round(0.75*(N_all))
N_test = N_all-N_train
diabetes_train <- diabetes[1:N_train,]
diabetes_test <- diabetes[N_train+1:N_all,]
classifier <- glm(X1~X6+X148+X72+X35+X0+X33.6+X0.627+X50,
                  family = 'binomial',
                  data = diabetes_train)
```

```{r confusion matrix}

predictions <- predict(classifier,newdata=diabetes_test, 
                       type = 'response')[1:N_test]
 
confusionMatrix(factor(round(predictions)),factor(diabetes_test['X1'][1:N_test,]))

```

```{r ROC}
diabetes_sensitivity <- sensitivity(factor(round(predictions)),factor(diabetes_test['X1'][1:N_test,]))
diabetes_specificity <- specificity(factor(round(predictions)),factor(diabetes_test['X1'][1:N_test,]))
diabetes_roc <- roc(round(predictions),diabetes_test['X1'][1:N_test,],
                    ci = TRUE, plot = TRUE, 
                    auc.polygon = TRUE, max.auc.polygon = TRUE, 
                    grid = TRUE, print.auc = TRUE, show.thres = TRUE)
diabetes.ci <- ci.se(diabetes_roc)
plot(diabetes.ci, type = 'shape')
plot(diabetes.ci, type = 'bars')
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

